---
title: "Tidytuesday: US Tornados"
format: html
editor: visual
---

## Quarto Tidytuesday: US Tornados

The data this week comes from NOAA's National Weather Service Storm Prediction Center Severe Weather Maps, Graphics, and Data Page. Thank you to Evan Gower for the suggestion!


```{r}
library(tidyverse)
library(quarto)
library(janitor)
library(lubridate)
library(scales)
library(plotly)
library(flextable)
```

## Read Data from Github Repo

```{r}
data <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-16/tornados.csv')
```
```{r}
## Count of Tornado's by Magnitude
data |> count(mag) |> ggplot(aes(x=mag,y=n)) + geom_col() +
  labs(title = "Count of Tornados by Magnitude",
       x="Magnitued",y="Count"        )
```

## # Plot 1: Scatterplot of Magnitude vs. Property Loss

```{r}

# plot1 <- ggplot(data, aes(x = mag, y = loss)) +
#   geom_point() +
#   xlab("Magnitude") +
#   ylab("Property Loss") +
#   ggtitle("Magnitude vs. Property Loss")

plot_ly(data, x = ~mag, y = ~loss, type = "scatter", mode = "markers") %>%
  layout(
    xaxis = list(title = "Magnitude"),
    yaxis = list(title = "Property Loss"),
    title = "Magnitude vs. Property Loss"
  )
```

## Plot 2: Bar Plot of Fatalities by Year

```{r}
# plot2 <- data %>%
#   group_by(yr) %>%
#   summarize(total_fatalities = sum(fat)) %>%
#   ggplot(aes(x = yr, y = total_fatalities)) +
#   geom_bar(stat = "identity") +
#   xlab("Year") +
#   ylab("Total Fatalities") +
#   ggtitle("Fatalities by Year")

data %>%
  group_by(yr) %>%
  summarize(total_fatalities = sum(fat)) %>%
  plot_ly(x = ~yr, y = ~total_fatalities, type = "bar") %>%
  layout(
    xaxis = list(title = "Year"),
    yaxis = list(title = "Total Fatalities"),
    title = "Fatalities by Year"
  )

```

## Plot 3: Line Plot of Tornado Length Over Time

```{r}
# plot3 <- data %>%
#   mutate(datetime_utc = as.POSIXct(datetime_utc)) %>%
#   ggplot(aes(x = datetime_utc, y = len)) +
#   geom_line() +
#   xlab("Date") +
#   ylab("Tornado Length (miles)") +
#   ggtitle("Tornado Length Over Time")

data %>%
  mutate(datetime_utc = as.POSIXct(datetime_utc)) %>%
  plot_ly(x = ~datetime_utc, y = ~len, type = "scatter", mode = "lines") %>%
  layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "Tornado Length (miles)"),
    title = "Tornado Length Over Time"
  )

```
```{r}
plot_ly(data, x = ~mag, y = ~loss, type = "bar", color = ~loss) %>%
  layout(
    xaxis = list(title = "Magnitude"),
    yaxis = list(title = "Property Loss"),
    title = "Magnitude vs. Property Loss"
  )
```

## Map: Scatterplot of Tornado Locations

```{r}
# library(leaflet)
# 
# # Create a leaflet map object
# map <- leaflet(data)
# 
# # Add markers for tornado locations
# addMarkers(
#   map,
#   lat = ~slat,
#   lng = ~slon,
#   popup = ~paste("Latitude: ", slat, "<br>Longitude: ", slon),
#   label = ~as.character(om)
# )

```

